Exploring environmental impacts of colonial contacts on Indigenous communities in northeastern Taiwan

### 2025-02-14 analysis/paper/paper.qmd
fixed grammar errors

Since deer remains was the major component of faunal assemblage at Kiwulan according to their high quantity and ubiquity, the frequency of skeletal elements were examined to explore any changes in processing decisions of meat portions between different phases. Deer bones were grouped into five body portions based on their nutritional value [@Madrigal2004], including head (cranium, mandible), axial (vertebrae, sternum, ribs), upper forelimb (scapula, humerus, ulna, and radius), upper hindlimb (femur, tibia, innominate), and lower limb (carpals, metacarpals, tarsals, metatarsals, astragalus, calcaneus, and phalanges). The result (@fig-deer-portion: A) shows that the meaty portion, including upper forelimb and upper hindlimb, especially upper forelimb, is the most frequent part throughout the cultural layers, followed by the less-meaty portions, such as lowerlimb and foot. The consistency of higher numbers of upperlimb and lowerlimb suggests that those elements should be transported together (e.g., @Stiner1991), or marrows from upperlimb were commonly utilized. It should be noted that the category of axial such as ribs were generally low across different phases, which might be caused by highly fragmentation, butchery practices or dietary preferences. A significant change is observed from the frequency of head bones that shows a few in CL6, increased and reach the highest number in CL4, and then decreased since then.

Considering varying frequencies of skeletal parts might be mediated by differential bone density, the relationship between the survivorship of deer skeletal elements and bone mineral density was tested by Pearson correlation. The values of bone survivorship was collected from @Lyman1994's research on deer, and calculated by dividing NISP by expected NISP, followed by multiplying 100%. The result (@fig-deer-portion: B) shows that there is no significant positive correlation ($r$ = `r survivor_cor[1]`, P = `r survivor_cor[2]`)) between skeletal part survivorship and bone density, suggesting that density-mediated bias has little impact on the taphonomic processes of deer bones.

>>>>

Since deer remain was the major component of faunal assemblage at Kiwulan according to their high quantity and ubiquity, the frequency of skeletal elements was examined to explore any changes in processing decisions of meat portions between different phases. Deer bones were grouped into five body portions based on their nutritional value [@Madrigal2004], including head (cranium, mandible), axial (vertebrae, sternum, ribs), upper forelimb (scapula, humerus, ulna, and radius), upper hindlimb (femur, tibia, innominate), and lower limb (carpals, metacarpals, tarsals, metatarsals, astragalus, calcaneus, and phalanges). The result (@fig-deer-portion: A) shows that the meaty portion, including upper forelimb and upper hindlimb, especially upper forelimb, is the most frequent part throughout the cultural layers, followed by the less-meaty portions, such as lower limb and foot. The consistency of higher numbers of upper limb and lower limb suggests that those elements should be transported together (e.g., @Stiner1991), or marrows from upper limb were commonly utilized. It should be noted that the frequencies of axial elements, such as ribs, were generally low across different phases, likely due to a high degree of fragmentation, butchery practices, or dietary preferences. A different pattern is observed in the frequency of head bones, which were a few in CL6, increased to the highest number in CL4, and then decreased thereafter.

Considering varying frequencies of skeletal parts might be mediated by differential bone density, the relationship between the survivorship of deer skeletal elements and bone mineral density was tested by Pearson correlation. The values of bone survivorship were collected from @Lyman1994's research on deer, and calculated by dividing NISP by expected NISP, followed by multiplying 100%. The result (@fig-deer-portion: B) shows that there is no significant positive correlation ($r$ = `r survivor_cor[1]`, P = `r survivor_cor[2]`)) between skeletal part survivorship and bone density, suggesting that density-mediated bias has little impact on the taphonomic processes of deer bones.
